---
title: "DON_data"
author: "Zachary Noel"
date: "12/12/2019"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r LIBRARIES}
library(plyr)
library(dplyr)
library(tidyr)
library(lme4)
library(emmeans)
library(cowplot)
library(ggpubr)
library(ggplot2)
library(multcomp)
```
## MYCOTOXIN DATA
```{r READ MYCOTOXIN DATA}
# DON data
DON <- read.csv("/Users/zacharynoel/odrive/WorkGoogleDrive/Post-doc/Wheat\ Biocontrol\ Project/R/WheatBiocontrol/Data/MycotoxinData.csv", na.strings = "na")
str(DON)

levels(as.factor(DON$Treatment))
```

Here we are going to analyse mycotoxin data that has been collected from wheat heads grown in the greenhouse inoculated with Fusarium graminearum or endophytes to protect against Fusarium graminearum. 

The experiment was performed by inoculating the heads with Fg and glycerol (positive control), Fg + endophytes, or glycerol (Negative control).

Therefore I will subset these data to only include those treatments. We will also take out the treatment with Fg + 37 + 40 since that treatment was only performed twice not three times like the rest of the treatmnts.
```{r}
DON_sub <- DON %>%
  subset(Treatment %in% c("NTC", "Fg", "Fg + 37", "Fg + 40", "Fg + 70")) 
```

### DON
Linear model DON - the model we are using is one where I am using the treatment as a fixed effect and the bio-rep (replication of the experiment over time) as a random effect. 
```{r LINEAR MODEL}
lm.DON <- lmer(DON ~ Treatment*Cultivar + (1|BioRep), data = DON_sub)
car::Anova(lm.DON)

plot(lm.DON) 
```
The model looks a little non-homogeneous but I think that is from the NTC and Glycerol control not having any DON and no variation. And the FG inoculated having alot of variation. 

```{r LSMEANS}
lsmeans.DON <- emmeans(lm.DON, ~Treatment) # estimate lsmeans of variety within siteXyear
Results_lsmeansEC <- emmeans::cld(lsmeans.DON, alpha = 0.05, adjust = "tuk", reversed = TRUE, details = TRUE) # contrast with Tukey ajustment
Results_lsmeansEC
```

We could also test this with a non-parametric Kruskal-Wallis test. 
```{r NON-PARAMETRIC}
kruskal.test(DON ~ Treatment, data = DON_sub)

pwwt <- pairwise.wilcox.test(DON_sub$DON, DON_sub$Treatment,
                 p.adjust.method = "bonferroni")

library(multcompView)
pwwt$p.value
library(reshape)
(a <- melt(pwwt$p.value))
a.cc  <-  na.omit(a)
a.pvals  <-  a.cc[, 3]
names(a.pvals)  <-  paste(a.cc[, 1], a.cc[, 2], sep="-")
a.pvals
multcompLetters(a.pvals)

melt(pwwt$p.value)

DON_sub %>% 
  group_by(Treatment) %>%
  summarize(avg = mean(DON, na.rm = T), 
            n = n(), 
            sd = sd(DON, na.rm = T), 
            se = sd/sqrt(n))
```
Seems like similar trends are shown to the parametric linear model. 

All endophytes alone were significantly different than the glycerol + Fg treatment, but not the combination of endophytes.

### 15A-DON

Linear model DON - the model we are using is one where I am using the treatment as a fixed effect and the bio-rep (replication of the experiment over time) as a random effect. 
```{r LINEAR MODEL}
lm.15ADON <- lmer(X15ADON ~ Treatment*Cultivar + (1|BioRep), data = DON_sub)
car::Anova(lm.15ADON)

plot(lm.15ADON) #looks a little non-homogeneous but I think that is from the NTC not having any DON and no variation 
```
The model also looks non-homogenous due to some larger variation between some treatments

```{r LSMEANS}
lsmeans.15ADON <- emmeans(lm.15ADON, ~Treatment) # estimate lsmeans of variety within siteXyear
Results_lsmeans15Adon <- emmeans::cld(lsmeans.15ADON, alpha = 0.05, adjust = "tuk", reversed = TRUE, details = TRUE) # contrast with Tukey ajustment
Results_lsmeans15Adon
```

We will also test this with a non-parametric Kruskal-Wallis test. 
```{r NON-PARAMETRIC}
kruskal.test(X15ADON ~ Treatment, data = DON_sub)

pwwt <- pairwise.wilcox.test(DON_sub$X15ADON, DON_sub$Treatment,
                 p.adjust.method = "bonferroni")

library(multcompView)
pwwt$p.value
library(reshape)
(a <- melt(pwwt$p.value))
a.cc  <-  na.omit(a)
a.pvals  <-  a.cc[, 3]
names(a.pvals)  <-  paste(a.cc[, 1], a.cc[, 2], sep="-")
a.pvals
multcompLetters(a.pvals)

melt(pwwt$p.value)

DON_sub %>% 
  group_by(Treatment) %>%
  summarize(avg = mean(X15ADON, na.rm = T), 
            n = n(), 
            sd = sd(X15ADON, na.rm = T), 
            se = sd/sqrt(n))
```
Seems like similar trends are shown to the parametric linear model. 
The endophytes reduced 15A DON compared to the Fg inoculated, but not quite to the level of the negative control

### Seed Mass

Linear model Seed Mass - the model we are using is one where I am using the treatment as a fixed effect and the bio-rep (replication of the experiment over time) as a random effect. 
```{r LINEAR MODEL}
lm.seedmass <- lmer(MassperSeed_mg ~ Treatment*Cultivar + (1|BioRep), data = DON_sub)
car::Anova(lm.seedmass, type = 3) # type 3 sum squares due to unequal sample sizes. 

plot(lm.seedmass) #looks much better than the DON data.
```
The model also looks non-homogenous due to some larger variation between some treatments

```{r LSMEANS}
lsmeans.seedmass <- emmeans(lm.seedmass, ~Treatment) # estimate lsmeans
Results_lsmeansseedmass <- multcomp::cld(lsmeans.seedmass, alpha = 0.05, adjust = "tuk", reversed = TRUE, details = TRUE) # contrast with Tukey ajustment
Results_lsmeansseedmass
```
Looks like there was not as strong of a violation of the non-homogenous variance across treatments so this can be analysed using a parametric test. 

```{r MEANS +/- SE}
DON_sub %>% 
  group_by(Treatment) %>%
  summarize(avg = mean(MassperSeed_mg), n = n(), 
              sd = sd(MassperSeed_mg), se = sd/sqrt(n))
```

### Plot

Lets make a plot
```{r PLOTS}
DON_sub$Treatment <- factor(DON_sub$Treatment,levels = c("NTC", "Fg", "Fg + 37", "Fg + 40", "Fg + 70"))

DON_total <- ggplot(DON_sub, aes(x = Treatment, y = DON)) +
  stat_summary(fun.y=mean,geom="bar", color = "black", width = 0.75) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
  geom_jitter(pch = 21, alpha = 0.6, width = 0.1) + 
  ylab("DON (ppm)") +
  stat_compare_means(label = "p.signif", method = "t.test",
                     ref.group = "Fg") +
  xlab("") +
theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_discrete(labels=c("Control", "Fg", "Fg + A. destruens #37", "Fg + F. commune #40", "Fg + F. oxysporum #70")) +
  facet_wrap(~Cultivar)
DON3A <- ggplot(DON_sub, aes(x = Treatment, y = X3ADON)) +
  stat_summary(fun.y=mean,geom="bar", color = "black", width = 0.75) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
  geom_jitter(pch = 21, alpha = 0.6, width = 0.1) + 
  ylab("3-ADON (ppm)") +
  xlab("") +
theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_discrete(labels=c("NTC", "Fg", "Fg + A. destruens #37", "Fg + F. commune #40", "Fg + F. oxysporum #70")) 
DON15A <- ggplot(DON_sub, aes(x = Treatment, y = X15ADON)) +
  stat_summary(fun.y=mean,geom="bar", color = "black", width = 0.75) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
  geom_jitter(pch = 21, alpha = 0.6, width = 0.1) + 
  ylab("15-ADON (ppm)") +
  xlab("") +
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_discrete(labels=c("Control", "Fg", "Fg + A. destruens #37", "Fg + F. commune #40", "Fg + F. oxysporum #70")) 

plot_grid(DON_total, DON15A, labels = "AUTO", ncol = 2)
```

## ROOT & SHOOT DRY WEIGHTS 
### Seed Mass

```{r READ DRY MASS DATA}
# Root and Shoot mass
root_shoot_mass <- read.csv("Data/Root_ShootDryWeight.csv", na.strings = "na")
str(root_shoot_mass)

# convert the dry mass data to mg 
root_shoot_mass$root_dry_weight_mg <- root_shoot_mass$root_dry_weight*1000
root_shoot_mass$shoot_dry_weight_mg <- root_shoot_mass$shoot_dry_weight*1000

# calculate plant mass
root_shoot_mass$total <- root_shoot_mass$root_dry_weight_mg + root_shoot_mass$shoot_dry_weight_mg

root_shoot_mass_sub <- root_shoot_mass %>%
  subset(Treatment %in% c("Fg", "Fg + 37", "Fg + 40", "Fg + 70"))
```

### ROOT MASS 
Linear model Root Mass - the model we are using is one where I am using the treatment as a fixed effect and the bio-rep (replication of the experiment over time) as a random effect. 
```{r LINEAR MODEL}
lm.rootmass <- lmer(root_dry_weight_mg ~ Treatment + (1|Biorep), data = root_shoot_mass_sub)
summary(lm.rootmass)
car::Anova(lm.rootmass, type = 3) # type 3 sum squares due to unequal sample sizes. 

plot(lm.rootmass) #looks much better than the DON data.
```
The model looks ok...not great but ok to continue I think

```{r LSMEANS}

lsmeans.rootmass <- emmeans(lm.rootmass, ~Treatment) # estimate lsmeans
Results_lsmeansrootmass <- multcomp::cld(lsmeans.rootmass, alpha = 0.05, adjust = "sidak", reversed = TRUE, details = TRUE) # contrast with Tukey ajustment
Results_lsmeansrootmass
```

### SHOOT MASS

Linear model Shoot Mass - the model we are using is one where I am using the treatment as a fixed effect and the bio-rep (replication of the experiment over time) as a random effect. 
```{r LINEAR MODEL}
lm.shootmass <- lm(shoot_dry_weight_mg ~ Treatment + Biorep, data = root_shoot_mass_sub)
car::Anova(lm.shootmass, type = 3) # type 3 sum squares due to unequal sample sizes. 

summary(lm.shootmass)

plot(lm.shootmass) #looks much better than the DON data.
```
The model looks ok

```{r LSMEANS}
lsmeans.shootmass <- emmeans(lm.shootmass, ~Treatment|Biorep) # estimate lsmeans
Results_lsmeansshootmass <- emmeans::cld(lsmeans.shootmass, alpha = 0.05, adjust = "tuk", reversed = TRUE, details = TRUE) # contrast with Tukey ajustment
Results_lsmeansshootmass
```

Linear model Total Plant Mass - the model we are using is one where I am using the treatment as a fixed effect and the bio-rep (replication of the experiment over time) as a random effect. 
```{r LINEAR MODEL}
lm.total <- lm(total ~ Treatment + Biorep, data = root_shoot_mass_sub)
car::Anova(lm.total, type = 3) # type 3 sum squares due to unequal sample sizes. 

plot(lm.shootmass) #looks much better than the DON data.
```

### Plot
```{r PLOTS}
root_shoot_mass$Treatment <- factor(root_shoot_mass$Treatment,levels = c("NTC", "Fg", "Fg + 37", "Fg + 40", "Fg + 70"))

RootMass <- ggplot(root_shoot_mass, aes(x = Treatment, y = root_dry_weight_mg)) +
  stat_summary(fun.y=mean,geom="bar", color = "black", width = 0.75) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
  geom_jitter(pch = 21, alpha = 0.6, width = 0.1) + 
  ylab("Root Mass (mg)") +
  xlab("") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ShootMass <- ggplot(root_shoot_mass, aes(x = Treatment, y = shoot_dry_weight_mg)) +
  stat_summary(fun.y=mean,geom="bar", color = "black", width = 0.75) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
  geom_jitter(pch = 21, alpha = 0.6, width = 0.1) + 
  ylab("Shoot Mass (mg)") +
  xlab("") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

SeedMass <- ggplot(DON_sub, aes(x = Treatment, y = MassperSeed_mg)) +
  stat_summary(fun.y=mean,geom="bar", color = "black", width = 0.75) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
  geom_jitter(pch = 21, alpha = 0.6, width = 0.1) + 
  ylab("Seed Mass (mg)") +
  xlab("") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+ 
  facet_wrap(~Cultivar) +
    stat_compare_means(label = "p.signif", method = "t.test",
                     ref.group = "Fg") 

ggplot(DON_sub, aes(x = MassperSeed_mg, y = log(DON), color = Treatment)) +
  geom_point() + 
  ylab("log DON") +
  xlab("Seed Mass (mg)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+ 
  facet_wrap(~Cultivar) +
    stat_compare_means(label = "p.signif", method = "t.test",
                     ref.group = "Fg") 

plot_grid(RootMass, ShootMass, SeedMass, labels = "AUTO", ncol = 1)
```

Put the cell means into a table form
```{r}
means <- root_shoot_mass %>% 
  group_by(Treatment) %>% 
  nest() %>%
  mutate(mean.rootmass = purrr::map(data,~mean(.$root_dry_weight_mg))) %>%
  mutate(SE.rootmass = purrr::map(data,~sd(.$root_dry_weight_mg)/sqrt(length(.$root_dry_weight_mg)))) %>%
  mutate(mean.shootmass = purrr::map(data,~mean(.$shoot_dry_weight_mg))) %>%
  mutate(SE.shootmass = purrr::map(data,~sd(.$shoot_dry_weight_mg)/sqrt(length(.$shoot_dry_weight_mg)))) %>%
  unnest(c(mean.rootmass, SE.rootmass, mean.shootmass, SE.shootmass))

means
```




